---
output: pdf_document
---
\textbf{\Large How well subjects did in bar bell lifts}

\Large Executive Summary  
bla bla bla...

\Large Data  
\normalsize Import Data from download cvs or from website link

```{r import_data, echo=TRUE}
if (!file.exists("pml-training.csv")) {
  download.file("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", 
                destfile = "pml-training.csv")
}
if (!file.exists("pml-testing.csv")) {
  download.file("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", 
                destfile = "pml-testing.csv")
}
```  

\normalsize Read in Data

```{r read_data, echo=TRUE}
pmlTraining <- read.csv("pml-training.csv", header = TRUE, sep = ",",na.strings=c("#DIV/0!"))
pmlTesting <- read.csv("pml-testing.csv", header = TRUE, sep = ",",na.strings=c("#DIV/0!"))
# View(pmlTraining)
pmlTraining[ pmlTraining == '' | pmlTraining == 'NA'] <- NA
```  

\normalsize Clean Data 

Reviewing the data we find three things we need to adjust
\begin{itemize}
\item NA values
\item Missing values, which we deall with by turning them into NA above
\item Error Expressions, which we handle on import via na.strings  
\end{itemize}

